Structuring College Access

### 2025-06-27 index.qmd
update rq1 graphs + tables

insert_rq1_table <- function(metro) {
  table_df <- readRDS(file.path(tables_dir, str_c('rq1_table_', metro, '.rds')))
  table_df[is.na(table_df)] <- NA_real_

@@

  kbl(table_df %>% select(-variable), escape = F, col.names = c('', rep(c('Mean', 'SD', 'P25', 'P50', 'P75'), 3))) %>% 
    pack_rows(index = headings_group, colnum = 1, label_row_css = '') %>% 
    add_header_above(header = c(' ' = 1, '1980' = 5, '2000' = 5, '2020' = 5))
}

>>>>

insert_rq1_table <- function(metro) {
  table_df <- readRDS(file.path(tables_dir, str_c('rq1_table_', metro, '.rds')))
  table_df$value[is.na(table_df$value)] <- NA_real_

  var_order <- c(
    '% White, non-Hispanic', '% Asian, non-Hispanic', '% Black, non-Hispanic', '% Hispanic',
    '% Two+, non-Hispanic', '% NHPI, non-Hispanic', '% AIAN, non-Hispanic',
    'Median income', 'Mean income', '% in poverty', '% with BA+'
  )

  race_df <- table_df %>% filter(statistic == 'sum', str_detect(variable, '^[^%]*Hispanic'), !str_detect(variable, 'API')) %>%
    group_by(eps_codename, year, statistic) %>% mutate(value = value / sum(value) * 100) %>% ungroup() %>% 
    mutate(
      variable = str_c('% ', variable),
      value = if_else((year == '1980' & str_detect(variable, 'Asian|NHPI|AIAN|Two')), NA_real_, value)
    )

  pov_df <- table_df %>% filter(statistic == 'sum', variable %in% c('Poverty', 'Not poverty')) %>%
    group_by(eps_codename, year, statistic) %>% mutate(value = value / sum(value) * 100) %>% ungroup() %>% 
    filter(variable == 'Poverty') %>% 
    mutate(variable = '% in poverty')

  edu_df <- table_df %>% filter(statistic == 'sum', variable %in% c('Less than HS', 'HS', 'Less than BA', 'BA+')) %>%
    group_by(eps_codename, year, statistic) %>% mutate(value = value / sum(value) * 100) %>% ungroup() %>% 
    filter(variable == 'BA+') %>% 
    mutate(variable = '% with BA+')

  table_df <- bind_rows(
    race_df, pov_df, edu_df,
    table_df %>% filter(variable %in% var_order, statistic != 'sum')
  ) %>% 
    mutate(
      statistic = factor(statistic, levels = c('sum', 'mean', 'sd', 'p25', 'p50', 'p75')),
      variable = factor(variable, levels = var_order),
      value = sprintf('%.1f', value)
    ) %>% 
    arrange(desc(year), statistic) %>% 
    pivot_wider(id_cols = c(eps_codename, variable), names_from = c(statistic, year), values_from = value) %>% 
    arrange(variable, desc(eps_codename))

@@

  kbl(table_df %>% select(-variable), escape = F, col.names = c('', rep(c('Wt. Mean', 'Mean', 'SD', 'P25', 'P50', 'P75'), 3))) %>% 
    pack_rows(index = headings_group, colnum = 1, label_row_css = '') %>% 
    add_header_above(header = c(' ' = 1, '1980' = 6, '2000' = 6, '2020' = 6))
}
